\documentclass[../thesis.tex]{subfiles}

\begin{document}

\chapter{Conclusion}
\label{chap:conclusion}

In \autoref{eq:fitResults}, we gave the best-fit oscillation parameters obtained when using the nominal IBD selection cuts. These values form the basis of our final result. However, in \autoref{sec:cutVaryJoint}, we assessed an additional systematic of 0.0006 ($1.9\times10^{-5}$~eV$^2$) on $\SinSq$ ($\Dmsqee$), based on the degree of variation observed when the IBD selection cuts were randomly fluctuated. We therefore add this systematic (in quadrature%
\footnote{In principle, this uncertainty may be partially correlated with the systematics encoded in the toy MC's covariance matrix (\autoref{sec:toymc}). If this correlation were complete, and the total statistical uncertainty were nonexistent, then it would be appropriate to add this uncertainty linearly, rather than in quadrature. However, given that the P17B data sample induces a statistical uncertainty that is nearly equal to the total systematic uncertainty \cite{p17b_osc}, and given that we have conservatively attributed the oscillation fit's cut-dependence entirely to systematics (thus ignoring its statistical component), the correlation between the cut-variation systematic and the nominal systematics is expected to be marginal, justifying the use of the quadrature sum. A fitter based on nuisance parameters, rather than on covariance matrices, may be better suited for studying any potential correlations between the cut-variation systematic and the nominal set of systematics.}%
) to the fitter's reported (statistical + systematic) uncertainty of 0.0029 ($6.9\times10^{-5}$~eV) from \autoref{eq:fitResults}, giving our final result:
\begin{equation}
  \SinSq = 0.0850 \pm 0.0030 \qquad \Dmsqee = (2.5010 \pm 0.0072)\times10^{-3}\;\text{eV}^2,
\end{equation}
We regard these uncertainties as comprehensive: They include the effects of statistics (from the toy MC's covariance matrix), signal/background systematics (again, from the toy MC), and the systematic uncertainty (added manually) arising from the cut-variation study. This work represents the most detailed assessment to-date of this final uncertainty, demonstrating that the Daya Bay experiment and oscillation analysis are of sufficiently high quality as to be largely insensitive to cut variations. In the worst case ($\Dmsqee$), the total uncertainty is increased by only 4\% when we account for cut variations. Nevertheless, it would be worthwhile for future Daya Bay results to be verified as stable against changes in IBD cuts; this work thus provides a foundation for further increasing the community's high confidence in the results of Daya Bay.

In closing, we have presented a detailed description of the Daya Bay oscillation analysis, including original work in such areas as the IBD selection, the $^9$Li and accidentals backgrounds, the efficiencies of various IBD cuts, and the synthesis of these efforts in our study of the effects of varying the IBD cuts. These results are built upon our lower-level work in such areas as calibration and data quality, along with the work of the many others who have helped to design, build, run, and analyze the experiment. Having provided additional evidence of the robustness of Daya Bay's analysis, it is our hope and expectation that confident knowledge in the values of $\SinSq$ and $\Dmsqee$ will help to enable the next generation of discoveries in neutrino physics. For the reader who has not yet had enough of neutrino physics, we explore some of these exciting possibilities in \autoref{sec:futureRelevance}.

% In summary, we have presented a detailed description of the Daya Bay oscillation analysis, going step-by-step from raw data to the final oscillation fit. Many individuals have contributed to the various steps in the analysis. Without their efforts, this work would not have been possible. The author's own contributions have focused primarily on calibration, reconstruction, data quality, and event selection. In particular, we have implemented an entirely new event selection framework, which was not only used for the IBD selection in this analysis, but which has also been adopted by at least one other analysis group for their own research. The high performance of our event selection enabled the analysis to be repeated for hundreds of variations of the IBD selection cuts. This allowed for a novel contribution to Daya Bay's results. Namely, we have shown two findings: First, that the best-fit oscillation parameters are stable under variations of the selection criteria, and second, that the error bars are also stable, implying that there is no compelling reason to modify the ad-hoc cuts that have been used in Daya Bay's publications. It is hoped that this work may bolster confidence in Daya Bay's measured oscillation parameters.

% Having measured $\theta_{13}$ (and $\Dmsqee$), we now conclude this thesis by discussing the significance of this measurement in the context of resolving a number of outstanding questions involving the neutrino and our Universe as a whole.


\end{document}